---
output: pdf_document
---
Assignment: Statistical Inference Course Project
==========================================

## Part 2: ToothGrowth data analyze.

### Overview  

Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package.

### TodoList

1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
4. State your conclusions and the assumptions needed for your conclusions.

### Part 0 - Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
set.seed(42)
library(ggplot2)
library(datasets)
data <- ToothGrowth
```

Some usefull plots

```{r}
ggplot(data = data, aes(x = dose, y = len, fill = supp))  +
  geom_bar(stat = "identity") +
  facet_grid(. ~ supp) +
  xlab("Dose in miligrams") +
  ylab("Tooth length") +
  guides(fill = guide_legend(title = "Supplement type"))
```

```{r}
ggplot(data = data, aes(x = as.factor(dose), y = len, fill = dose)) +
  geom_boxplot(aes(fill = dose)) +
  xlab("Dose in miligrams") +
  ylab("Tooth length") +
  guides(fill = guide_legend(title = "Dose"))
```

```{r}
ggplot(data = data, aes(x = supp, y = len, fill = dose))  +
  geom_boxplot(aes(fill = supp)) +
  xlab("Supplement type") +
  ylab("Tooth length") +
  guides(fill = guide_legend(title = "Supplement type"))
```

#### Part 1 - Provide a basic summary of the data

```{r}
summary(data)
```  

```{r}
str(data)
```  

```{r}
table(data$dose, data$supp)
```  



#### Part 2 - Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

Since the sample size is small, let's use T distribution.

```{r}
t.test(len ~ supp, data = data)
```  

Null hypothesis can not be rejected as confindence intervals contain zero and p-value is bigger than the 5% significance level.


Let's also compare the differences between the the different dosis level, since bigger dosis may yield contradicting evidence.

```{r}
t.test(data$len, data$dose)
```  

Comparing the difference between the two supplements yields convincing evidence to reject the null hypothesis, since the p-value aproximates to 0, and is thus substancially smaller in comparison to the significance level.

#### Part 3 - State your conclusions and the assumptions needed for your conclusions.

1. There is no convincing evidence that there is a difference between the two type of supplements based on the existing datasets and T statistics.
2. There is convincing evidence that there is a difference between the dosis level, and the growth.
